---
title: "Customer Retention Project"
output: html_notebook
---

### Install Required Libraries
```{r}
# Uncomment respective line for which package is missing.
#install.packages("ggplot2")
#install.packages("tidyr")
#install.packages("PerformanceAnalytics") 
```

### Load Required Libraries
```{r}
library(ggplot2)
library(tidyr)
library(PerformanceAnalytics) 
library(dplyr)
library(caret)
library(pROC)
```

### Datasets

### Load Datasets
```{r}
implementation_data <- "Implementation_Data.rda"
load(implementation_data)

model_building_data <- "Model_Building_Data.rda"
load(model_building_data)
```

### Implementation Dataset Quick View
```{r}
head(Implementation_Data)
str(Implementation_Data)
```

### Model Building Dataset Quick View
```{r}
head(Model_Building_Data)
str(Model_Building_Data)
```

# 1. Data Preparation & Exploratory Data Analysis
**Data preparation** is cleaning and organizing data, while **Exploratory Data Analysis (EDA)** helps explore patterns and trends using summaries and visuals.

## 1.1 Data Preparation
### Datatypes
**Status** is loaded as *chr* and having `r length(unique(Model_Building_Data$Status))` values only so it can be converted into *factor*
```{r}
Model_Building_Data$Status <- as.factor(Model_Building_Data$Status)
```

**SeniorCitizen** is loaded as an *integer* it can be converted into a *factor* for better machine learning analysis.
```{r}
Model_Building_Data$SeniorCitizen <- as.factor(Model_Building_Data$SeniorCitizen)
```

### Missing Values
There are multiple ways to handle missing values:
* Removing -> If missing values are very few <5% of data.
* Mean -> If normally distributed.
* Median -> If data is skewed.
* Regression -> Many missing values.

Let's check missing values for all the columns:
```{r}
sapply(Model_Building_Data, function(x) sum(is.na(x)))
```
We have 11 missing values for **TotalCharges** column which is very small less (<5% of data) so we can remove missing records.
Let's check the histogram and analyze the distribution of the data.
```{r}
hist(Model_Building_Data$TotalCharges, 
     breaks = 30, 
     main = "Distribution of Total Charges", 
     xlab = "Total Charges ($)", 
     ylab = "Customers", 
     col = "skyblue")

```
*TotalCharges* is right skewed, so we can use median in order to update missing records.
```{r}
Model_Building_Data$TotalCharges[is.na(Model_Building_Data$TotalCharges)] <- median(Model_Building_Data$TotalCharges, na.rm = TRUE)
```

Now cross check the whether missing values are handled or not
```{r}
sapply(Model_Building_Data, function(x) sum(is.na(x)))
```

## 1.2 Exploratory Data Analysis
TODO: Repeat for all Factor columns (Option 2) try one
```{r}
ggplot(Model_Building_Data, aes(x = Partner)) +
  geom_bar(fill = "skyblue", color="black", alpha = 0.6) +
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.6, size = 3) +
  theme_minimal(base_size = 12) +
  labs(title = "Bar Plot of Partner", x = "Partner", y = "Customers")
```

```{r}
ggplot(Model_Building_Data, aes(x = Dependents)) +
  geom_bar(fill = "skyblue", color="black", alpha = 0.6) +
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.6, size = 3) +
  theme_minimal(base_size = 12) +
  labs(title = "Bar Plot of Dependents", x = "Dependents", y = "Customers")
```

```{r}
ggplot(Model_Building_Data, aes(x = PhoneService)) +
  geom_bar(fill = "skyblue", color="black", alpha = 0.6) +
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.6, size = 3) +
  theme_minimal(base_size = 12) +
  labs(title = "Bar Plot of Phoneservice", x = "Phoneservice", y = "Customers")
```

```{r}
ggplot(Model_Building_Data, aes(x = MultipleLines)) +
  geom_bar(fill = "skyblue", color="black", alpha = 0.6) +
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.6, size = 3) +
  theme_minimal(base_size = 12) +
  labs(title = "Bar Plot of Multiplelines", x = "Multiplelines", y = "Customers")
```

```{r}
ggplot(Model_Building_Data, aes(x = InternetService)) +
  geom_bar(fill = "skyblue", color="black", alpha = 0.6) +
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.6, size = 3) +
  theme_minimal(base_size = 12) +
  labs(title = "Bar Plot of Internetservice", x = "Internetservice", y = "Customers")
```

```{r}
ggplot(Model_Building_Data, aes(x = OnlineSecurity)) +
  geom_bar(fill = "skyblue", color="black", alpha = 0.6) +
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.6, size = 3) +
  theme_minimal(base_size = 12) +
  labs(title = "Bar Plot of Onlinesecurity", x = "Onlinesecurity", y = "Customers")
```

```{r}
ggplot(Model_Building_Data, aes(x = DeviceProtection)) +
  geom_bar(fill = "skyblue", color="black", alpha = 0.6) +
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.6, size = 3) +
  theme_minimal(base_size = 12) +
  labs(title = "Bar Plot of Deviceprotection", x = "Deviceprotection", y = "Customers")
```

```{r}
ggplot(Model_Building_Data, aes(x = Contract)) +
  geom_bar(fill = "skyblue", color="black", alpha = 0.6) +
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.6, size = 3) +
  theme_minimal(base_size = 12) +
  labs(title = "Bar Plot of Contract", x = "Contract", y = "Customers")
```

```{r}
ggplot(Model_Building_Data, aes(x = StreamingMovies)) +
  geom_bar(fill = "skyblue", color="black", alpha = 0.6) +
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.6, size = 3) +
  theme_minimal(base_size = 12) +
  labs(title = "Bar Plot of Streamingmovies", x = "Streamingmovies", y = "Customers")
```

```{r}
ggplot(Model_Building_Data, aes(x = PaperlessBilling)) +
  geom_bar(fill = "skyblue", color="black", alpha = 0.6) +
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.6, size = 3) +
  theme_minimal(base_size = 12) +
  labs(title = "Bar Plot of Paperlessbilling", x = "Paperlessbilling", y = "Customers")
```

```{r}
ggplot(Model_Building_Data, aes(x = PaymentMethod)) +
  geom_bar(fill = "skyblue", color="black", alpha = 0.6) +
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.6, size = 3) +
  theme_minimal(base_size = 12) +
  labs(title = "Bar Plot of PaymentMethod", x = "PaymentMethod", y = "Customers")
```

```{r}
ggplot(Model_Building_Data, aes(x = Status)) +
  geom_bar(fill = "skyblue", color="black", alpha = 0.6) +
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.6, size = 3) +
  theme_minimal(base_size = 12) +
  labs(title = "Bar Plot of Status", x = "Status", y = "Customers")
```

TODO: Explain bar plot and your findings.

TODO: Repeat for all numeric columns and adjust bins as required
```{r}
ggplot(Model_Building_Data, aes(x = Tenure)) +
  geom_histogram(binwidth = 10, boundary = 0, fill = "skyblue", color = "black", alpha = 0.7) +
  labs(title = "Distribution of Tenure", x = "Tenure (Bin=10)", y = "Customers") +
  theme_minimal()
```

```{r}
ggplot(Model_Building_Data, aes(x = MonthlyCharges)) +
  geom_histogram(binwidth = 10, boundary = 0, fill = "skyblue", color = "black", alpha = 0.7) +
  labs(title = "Distribution of MonthlyCharges", x = "Monthlycharges (Bin=10)", y = "Customers") +
  theme_minimal()
```

```{r}
ggplot(Model_Building_Data, aes(x = TotalCharges)) +
  geom_histogram(binwidth = 1000, boundary = 0, fill = "skyblue", color = "black", alpha = 0.7) +
  labs(title = "Distribution of TotalCharges", x = "Totalcharges (Bin=10)", y = "Customers") +
  theme_minimal()
```
TODO: Explain histogram plot and your findings.

### Correlation of Customer
```{r}
Model_Building_Data.numeric <- Model_Building_Data %>%
  select(Tenure,MonthlyCharges,TotalCharges)
chart.Correlation(Model_Building_Data.numeric , histogram=TRUE, pch=19)
```


### Spliting the data
```{r}
train <- sample(nrow(Model_Building_Data), size = 0.8 * nrow(Model_Building_Data))

model.train <- Model_Building_Data[train,]
model.test <- Model_Building_Data[-train,]
```

# 2. Logistic Regression

### 2.1 bilding logistic regression

```{r}
logit_model <- glm(Status ~ ., data = model.train, family = binomial)
summary(logit_model)
```

```{r}
model.test.predicted_prob <- predict(logit_model, newdata = model.test, type = "response")
roc_curve <- roc(model.test$Status, model.test.predicted_prob)
auc_value <- auc(roc_curve)
auc_value
```

```{r}
plot(roc_curve, col = "blue", main = paste("ROC Curve (AUC =", round(auc_value, 4), ")"))
```

```{r}

 Prediction_Classes <- ifelse(model.test.predicted_prob > 0.5, "Left", "Current")
  conf_matrix <- table(Prediction_Classes, model.test$Status)
 
  accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix) * 100
accuracy
```

```{r}
 Prediction_Classes <- ifelse(model.test.predicted_prob > 0.3, "Left", "Current")
  conf_matrix <- table(Prediction_Classes, model.test$Status)
 
  accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix) * 100
accuracy
```

```{r}
 Prediction_Classes <- ifelse(model.test.predicted_prob > 0.7, "Left", "Current")
  conf_matrix <- table(Prediction_Classes, model.test$Status)
 
  accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix) * 100
accuracy
```
